---
author: "Camille Ross"
date: "8/17/2020"
output:
  html_document:
    toc: true
    toc_float: true
params:
  set_title: "v0.2.3"
  fp_out: "./calanus_for_whales/Versions"
  species: "cfin"
  version: "v0.2.3"
  env_covars: c("sst")
  threshold: "40,000"
  years: "2000:2017"
  datasets: "CPR"
  biomod_dataset: "Ecomon"
title: "`r params$set_title`"
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_libraries, echo=FALSE, include = FALSE, message=FALSE}
library(png)
library(grid)
library(gridExtra)
```

## Summary

Model ``r params$version`` was created using ``r params$env_covars`` for ``r params$species``.  The models were averaged into climatologies with one climatology per month.  Evaluations were compiled from the climatological averages and plotted.  Finally, the study area was divided up into three regions, the Mid-Atlantic Bight (MAB), George's Bank (GBK), and the Gulf of Maine (GOM).  Actual versus predicted abundance values were plotted for each region.  The mgvc GAMs and the gbm BRTs were run using the ``r params$datasets`` dataset(s).  The Biomod2 models were run using the ``r params$biomod_dataset`` dataset(s).  If the model is an anomaly, all datasets are used.  

## Biomod: right whale feeding threshold models

### Climatologies

#### Biomod Ensemble Climatology

The ensemble models were created using the biomod2 package.  The ensembles consist of BRTs, GAMs, and random forests (RFs).  The ensembles were used to model the right whale feeding threshold, with any abundance greater than ``r params$threshold`` ``r params$species`` per $m^2$ counted as a presence and anything below that threshold counted as an absence.

```{r display_ensemble, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Plots")

# Load images
ensemble1 <- readPNG(file.path(fp, "ensemble_proj_1.png"))
ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2.png"))
ensemble3 <- readPNG(file.path(fp, "ensemble_proj_3.png"))
ensemble4 <- readPNG(file.path(fp, "ensemble_proj_4.png"))
ensemble5 <- readPNG(file.path(fp, "ensemble_proj_5.png"))
ensemble6 <- readPNG(file.path(fp, "ensemble_proj_6.png"))
ensemble7 <- readPNG(file.path(fp, "ensemble_proj_7.png"))
ensemble8 <- readPNG(file.path(fp, "ensemble_proj_8.png"))
ensemble9 <- readPNG(file.path(fp, "ensemble_proj_9.png"))
ensemble10 <- readPNG(file.path(fp, "ensemble_proj_10.png"))
ensemble11 <- readPNG(file.path(fp, "ensemble_proj_11.png"))
ensemble12 <- readPNG(file.path(fp, "ensemble_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
             rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
             rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
             rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
             ncol=3)


```

**Figure 1.** Monthly climatological ensemble projections of GAMs, BRTs, and random forests (RFs).  The climatology was created by averaging together the projections from 2000 to 2017.

#### GAM Climatology

The GAM models created with biomod2 were used to model the right whale feeding threshold, with any abundance greater than ``r params$threshold`` ``r params$species`` per $m^2$ counted as a presence and anything below that threshold counted as an absence.  

```{r display_biomod_gam, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Plots")

# Load images
gam1 <- readPNG(file.path(fp, "gam_proj_1.png"))
gam2 <- readPNG(file.path(fp, "gam_proj_2.png"))
gam3 <- readPNG(file.path(fp, "gam_proj_3.png"))
gam4 <- readPNG(file.path(fp, "gam_proj_4.png"))
gam5 <- readPNG(file.path(fp, "gam_proj_5.png"))
gam6 <- readPNG(file.path(fp, "gam_proj_6.png"))
gam7 <- readPNG(file.path(fp, "gam_proj_7.png"))
gam8 <- readPNG(file.path(fp, "gam_proj_8.png"))
gam9 <- readPNG(file.path(fp, "gam_proj_9.png"))
gam10 <- readPNG(file.path(fp, "gam_proj_10.png"))
gam11 <- readPNG(file.path(fp, "gam_proj_11.png"))
gam12 <- readPNG(file.path(fp, "gam_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(gam1), rasterGrob(gam2), rasterGrob(gam3),
             rasterGrob(gam4), rasterGrob(gam5), rasterGrob(gam6),
             rasterGrob(gam7), rasterGrob(gam8), rasterGrob(gam9),
             rasterGrob(gam10), rasterGrob(gam11), rasterGrob(gam12),
             ncol=3)


```

**Figure 2.** Monthly climatological GAM projections produced using Biomod2.  The climatology was created by averaging together the projections from 2000 to 2017.

#### BRT Climatology

The BRT models created with biomod2 were used to model the right whale feeding threshold, with any abundance greater than ``r params$threshold`` ``r params$species`` per $m^2$ counted as a presence and anything below that threshold counted as an absence.  

```{r display_biomod_brt, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Plots")

# Load images
brt1 <- readPNG(file.path(fp, "brt_proj_1.png"))
brt2 <- readPNG(file.path(fp, "brt_proj_2.png"))
brt3 <- readPNG(file.path(fp, "brt_proj_3.png"))
brt4 <- readPNG(file.path(fp, "brt_proj_4.png"))
brt5 <- readPNG(file.path(fp, "brt_proj_5.png"))
brt6 <- readPNG(file.path(fp, "brt_proj_6.png"))
brt7 <- readPNG(file.path(fp, "brt_proj_7.png"))
brt8 <- readPNG(file.path(fp, "brt_proj_8.png"))
brt9 <- readPNG(file.path(fp, "brt_proj_9.png"))
brt10 <- readPNG(file.path(fp, "brt_proj_10.png"))
brt11 <- readPNG(file.path(fp, "brt_proj_11.png"))
brt12 <- readPNG(file.path(fp, "brt_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(brt1), rasterGrob(brt2), rasterGrob(brt3),
             rasterGrob(brt4), rasterGrob(brt5), rasterGrob(brt6),
             rasterGrob(brt7), rasterGrob(brt8), rasterGrob(brt9),
             rasterGrob(brt10), rasterGrob(brt11), rasterGrob(brt12),
             ncol=3)


```

**Figure 3.** Monthly climatological BRT projections produced using Biomod2.  The climatology was created by averaging together the projections from 2000 to 2017.

#### RF Climatology

The RF models created with biomod2 were used to model the right whale feeding threshold, with any abundance greater than ``r params$threshold`` ``r params$species`` per $m^2$ counted as a presence and anything below that threshold counted as an absence.  

```{r display_biomod_rf, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Plots")

# Load images
rf1 <- readPNG(file.path(fp, "rf_proj_1.png"))
rf2 <- readPNG(file.path(fp, "rf_proj_2.png"))
rf3 <- readPNG(file.path(fp, "rf_proj_3.png"))
rf4 <- readPNG(file.path(fp, "rf_proj_4.png"))
rf5 <- readPNG(file.path(fp, "rf_proj_5.png"))
rf6 <- readPNG(file.path(fp, "rf_proj_6.png"))
rf7 <- readPNG(file.path(fp, "rf_proj_7.png"))
rf8 <- readPNG(file.path(fp, "rf_proj_8.png"))
rf9 <- readPNG(file.path(fp, "rf_proj_9.png"))
rf10 <- readPNG(file.path(fp, "rf_proj_10.png"))
rf11 <- readPNG(file.path(fp, "rf_proj_11.png"))
rf12 <- readPNG(file.path(fp, "rf_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(rf1), rasterGrob(rf2), rasterGrob(rf3),
             rasterGrob(rf4), rasterGrob(rf5), rasterGrob(rf6),
             rasterGrob(rf7), rasterGrob(rf8), rasterGrob(rf9),
             rasterGrob(rf10), rasterGrob(rf11), rasterGrob(rf12),
             ncol=3)


```

**Figure 4.** Monthly climatological RF projections produced using Biomod2.  The climatology was created by averaging together the projections from 2000 to 2017.

### Monthly ensemble projections

Monthly ensemble Biomod2 projections are displayed below for the months of May, June, July, August, and September.

#### April

```{r display_biomod_april, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")
files <- list.files(fp)

# Load images
if ("ensemble_proj_2002_4.png" %in% files) {
  ensemble1 <- readPNG(file.path(fp, "ensemble_proj_2000_4.png"))
  ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2001_4.png"))
  ensemble3 <- readPNG(file.path(fp, "ensemble_proj_2002_4.png"))
  ensemble4 <- readPNG(file.path(fp, "ensemble_proj_2003_4.png"))
  ensemble5 <- readPNG(file.path(fp, "ensemble_proj_2004_4.png"))
  ensemble6 <- readPNG(file.path(fp, "ensemble_proj_2005_4.png"))
  ensemble7 <- readPNG(file.path(fp, "ensemble_proj_2006_4.png"))
  ensemble8 <- readPNG(file.path(fp, "ensemble_proj_2007_4.png"))
  ensemble9 <- readPNG(file.path(fp, "ensemble_proj_2008_4.png"))
  ensemble10 <- readPNG(file.path(fp, "ensemble_proj_2009_4.png"))
  ensemble11 <- readPNG(file.path(fp, "ensemble_proj_2010_4.png"))
  ensemble12 <- readPNG(file.path(fp, "ensemble_proj_2011_4.png"))
  ensemble13 <- readPNG(file.path(fp, "ensemble_proj_2012_4.png"))
  ensemble14 <- readPNG(file.path(fp, "ensemble_proj_2013_4.png"))
  ensemble15 <- readPNG(file.path(fp, "ensemble_proj_2014_4.png"))
  ensemble16 <- readPNG(file.path(fp, "ensemble_proj_2015_4.png"))
  ensemble17 <- readPNG(file.path(fp, "ensemble_proj_2016_4.png"))
  ensemble18 <- readPNG(file.path(fp, "ensemble_proj_2017_4.png"))
  
  # Arrange grid
  grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
               rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
               rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
               rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
               rasterGrob(ensemble13), rasterGrob(ensemble14), rasterGrob(ensemble15),
               rasterGrob(ensemble16), rasterGrob(ensemble17),  rasterGrob(ensemble18),
               ncol=3)
}

```

**Figure 5.** Ensemble projections for the month of April over ``r params$years``.

#### May

```{r display_biomod_may, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load images
ensemble1 <- readPNG(file.path(fp, "ensemble_proj_2000_5.png"))
ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2001_5.png"))
ensemble3 <- readPNG(file.path(fp, "ensemble_proj_2002_5.png"))
ensemble4 <- readPNG(file.path(fp, "ensemble_proj_2003_5.png"))
ensemble5 <- readPNG(file.path(fp, "ensemble_proj_2004_5.png"))
ensemble6 <- readPNG(file.path(fp, "ensemble_proj_2005_5.png"))
ensemble7 <- readPNG(file.path(fp, "ensemble_proj_2006_5.png"))
ensemble8 <- readPNG(file.path(fp, "ensemble_proj_2007_5.png"))
ensemble9 <- readPNG(file.path(fp, "ensemble_proj_2008_5.png"))
ensemble10 <- readPNG(file.path(fp, "ensemble_proj_2009_5.png"))
ensemble11 <- readPNG(file.path(fp, "ensemble_proj_2010_5.png"))
ensemble12 <- readPNG(file.path(fp, "ensemble_proj_2011_5.png"))
ensemble13 <- readPNG(file.path(fp, "ensemble_proj_2012_5.png"))
ensemble14 <- readPNG(file.path(fp, "ensemble_proj_2013_5.png"))
ensemble15 <- readPNG(file.path(fp, "ensemble_proj_2014_5.png"))
ensemble16 <- readPNG(file.path(fp, "ensemble_proj_2015_5.png"))
ensemble17 <- readPNG(file.path(fp, "ensemble_proj_2016_5.png"))
ensemble18 <- readPNG(file.path(fp, "ensemble_proj_2017_5.png"))

# Arrange grid
grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
             rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
             rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
             rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
             rasterGrob(ensemble13), rasterGrob(ensemble14), rasterGrob(ensemble15),
             rasterGrob(ensemble16), rasterGrob(ensemble17),  rasterGrob(ensemble18),
             ncol=3)
```

**Figure 6.** Ensemble projections for the month of May over ``r params$years``.

#### June

```{r display_biomod_june, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load images
ensemble1 <- readPNG(file.path(fp, "ensemble_proj_2000_6.png"))
ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2001_6.png"))
ensemble3 <- readPNG(file.path(fp, "ensemble_proj_2002_6.png"))
ensemble4 <- readPNG(file.path(fp, "ensemble_proj_2003_6.png"))
ensemble5 <- readPNG(file.path(fp, "ensemble_proj_2004_6.png"))
ensemble6 <- readPNG(file.path(fp, "ensemble_proj_2005_6.png"))
ensemble7 <- readPNG(file.path(fp, "ensemble_proj_2006_6.png"))
ensemble8 <- readPNG(file.path(fp, "ensemble_proj_2007_6.png"))
ensemble9 <- readPNG(file.path(fp, "ensemble_proj_2008_6.png"))
ensemble10 <- readPNG(file.path(fp, "ensemble_proj_2009_6.png"))
ensemble11 <- readPNG(file.path(fp, "ensemble_proj_2010_6.png"))
ensemble12 <- readPNG(file.path(fp, "ensemble_proj_2011_6.png"))
ensemble13 <- readPNG(file.path(fp, "ensemble_proj_2012_6.png"))
ensemble14 <- readPNG(file.path(fp, "ensemble_proj_2013_6.png"))
ensemble15 <- readPNG(file.path(fp, "ensemble_proj_2014_6.png"))
ensemble16 <- readPNG(file.path(fp, "ensemble_proj_2015_6.png"))
ensemble17 <- readPNG(file.path(fp, "ensemble_proj_2016_6.png"))
ensemble18 <- readPNG(file.path(fp, "ensemble_proj_2017_6.png"))

# Arrange grid
grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
             rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
             rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
             rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
             rasterGrob(ensemble13), rasterGrob(ensemble14), rasterGrob(ensemble15),
             rasterGrob(ensemble16), rasterGrob(ensemble17),  rasterGrob(ensemble18),
             ncol=3)
```

**Figure 7.** Ensemble projections for the month of June over ``r params$years``.

#### August

```{r display_biomod_august, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load images
ensemble1 <- readPNG(file.path(fp, "ensemble_proj_2000_8.png"))
ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2001_8.png"))
ensemble3 <- readPNG(file.path(fp, "ensemble_proj_2002_8.png"))
ensemble4 <- readPNG(file.path(fp, "ensemble_proj_2003_8.png"))
ensemble5 <- readPNG(file.path(fp, "ensemble_proj_2004_8.png"))
ensemble6 <- readPNG(file.path(fp, "ensemble_proj_2005_8.png"))
ensemble7 <- readPNG(file.path(fp, "ensemble_proj_2006_8.png"))
ensemble8 <- readPNG(file.path(fp, "ensemble_proj_2007_8.png"))
ensemble9 <- readPNG(file.path(fp, "ensemble_proj_2008_8.png"))
ensemble10 <- readPNG(file.path(fp, "ensemble_proj_2009_8.png"))
ensemble11 <- readPNG(file.path(fp, "ensemble_proj_2010_8.png"))
ensemble12 <- readPNG(file.path(fp, "ensemble_proj_2011_8.png"))
ensemble13 <- readPNG(file.path(fp, "ensemble_proj_2012_8.png"))
ensemble14 <- readPNG(file.path(fp, "ensemble_proj_2013_8.png"))
ensemble15 <- readPNG(file.path(fp, "ensemble_proj_2014_8.png"))
ensemble16 <- readPNG(file.path(fp, "ensemble_proj_2015_8.png"))
ensemble17 <- readPNG(file.path(fp, "ensemble_proj_2016_8.png"))
ensemble18 <- readPNG(file.path(fp, "ensemble_proj_2017_8.png"))

# Arrange grid
grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
             rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
             rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
             rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
             rasterGrob(ensemble13), rasterGrob(ensemble14), rasterGrob(ensemble15),
             rasterGrob(ensemble16), rasterGrob(ensemble17),  rasterGrob(ensemble18),
             ncol=3)
```

**Figure 8.** Ensemble projections for the month of August over ``r params$years``.

#### September

```{r display_biomod_september, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load images
  ensemble1 <- readPNG(file.path(fp, "ensemble_proj_2000_9.png"))
  ensemble2 <- readPNG(file.path(fp, "ensemble_proj_2001_9.png"))
  ensemble3 <- readPNG(file.path(fp, "ensemble_proj_2002_9.png"))
  ensemble4 <- readPNG(file.path(fp, "ensemble_proj_2003_9.png"))
  ensemble5 <- readPNG(file.path(fp, "ensemble_proj_2004_9.png"))
  ensemble6 <- readPNG(file.path(fp, "ensemble_proj_2005_9.png"))
  ensemble7 <- readPNG(file.path(fp, "ensemble_proj_2006_9.png"))
  ensemble8 <- readPNG(file.path(fp, "ensemble_proj_2007_9.png"))
  ensemble9 <- readPNG(file.path(fp, "ensemble_proj_2008_9.png"))
  ensemble10 <- readPNG(file.path(fp, "ensemble_proj_2009_9.png"))
  ensemble11 <- readPNG(file.path(fp, "ensemble_proj_2010_9.png"))
  ensemble12 <- readPNG(file.path(fp, "ensemble_proj_2011_9.png"))
  ensemble13 <- readPNG(file.path(fp, "ensemble_proj_2012_9.png"))
  ensemble14 <- readPNG(file.path(fp, "ensemble_proj_2013_9.png"))
  ensemble15 <- readPNG(file.path(fp, "ensemble_proj_2014_9.png"))
  ensemble16 <- readPNG(file.path(fp, "ensemble_proj_2015_9.png"))
  ensemble17 <- readPNG(file.path(fp, "ensemble_proj_2016_9.png"))
  ensemble18 <- readPNG(file.path(fp, "ensemble_proj_2017_9.png"))
  
  # Arrange grid
  grid.arrange(rasterGrob(ensemble1), rasterGrob(ensemble2), rasterGrob(ensemble3),
               rasterGrob(ensemble4), rasterGrob(ensemble5), rasterGrob(ensemble6),
               rasterGrob(ensemble7), rasterGrob(ensemble8), rasterGrob(ensemble9),
               rasterGrob(ensemble10), rasterGrob(ensemble11), rasterGrob(ensemble12),
               rasterGrob(ensemble13), rasterGrob(ensemble14), rasterGrob(ensemble15),
               rasterGrob(ensemble16), rasterGrob(ensemble17),  rasterGrob(ensemble18),
               ncol=3)

```

**Figure 9.** Ensemble projections for the month of September over ``r params$years``.


### Evaluations

Evaluation metrics differed based on the metrics available in each modeling package and compatible with each model object.  For the mgcv GAMs, Aikaike's Information Criterion (AIC), the root mean squared error (RMSE), and the R squared (RSQ) value when comparing the actual and predicted abundances were computed.  For the BRTs produced using the gbm package, RMSE and RSQ were computed.  For the biomod2 ensembles and GAMs, the area under the receiver operator characteristic curve (AUC) and the true skill statistic (TSS) were computed.  

#### Ensemble evaluations

```{r ensemble_evals, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Evals")

# Load evals
auc <- readPNG(file.path(fp, "ensemble_auc.png"))
tss <- readPNG(file.path(fp, "ensemble_tss.png"))

# Arrange grid
grid.arrange(rasterGrob(auc), rasterGrob(tss),
             ncol=2)
```

**Figure 10.** Biomod ensemble evaluations on a monthly time scale using a.) AUC and b.) TSS

#### GAM evaluations

```{r gam_biomod_evals, echo=FALSE}
# Load evals
auc <- readPNG(file.path(fp, "gam_auc.png"))
tss <- readPNG(file.path(fp, "gam_tss.png"))

# Arrange grid
grid.arrange(rasterGrob(auc), rasterGrob(tss),
             ncol=2)
```

**Figure 11.** Biomod GAM evaluations on a monthly time scale using a.) AUC and b.) TSS

#### BRT evaluations

```{r brt_biomod_evals, echo=FALSE}
# Load evals
auc <- readPNG(file.path(fp, "brt_auc.png"))
tss <- readPNG(file.path(fp, "brt_tss.png"))

# Arrange grid
grid.arrange(rasterGrob(auc), rasterGrob(tss),
             ncol=2)
```

**Figure 12.** Biomod BRT evaluations on a monthly time scale using a.) AUC and b.) TSS


#### RF evaluations

```{r rf_biomod_evals, echo=FALSE}
# Load evals
auc <- readPNG(file.path(fp, "rf_auc.png"))
tss <- readPNG(file.path(fp, "rf_tss.png"))

# Arrange grid
grid.arrange(rasterGrob(auc), rasterGrob(tss),
             ncol=2)
```

**Figure 13.** Biomod RF evaluations on a monthly time scale using a.) AUC and b.) TSS

### Variable contribution

#### GAM variable contribution

```{r gam_biomod_var_cont, echo=FALSE}
# Load evals
var_cont <- readPNG(file.path(fp, "gam_var_cont.png"))

# Arrange grid
grid.arrange(rasterGrob(var_cont),
             ncol = 1)
```

**Figure 14.** Biomod GAM variable contributions on a monthly time scale.

#### BRT variable contribution

```{r brt_biomod_var_cont, echo=FALSE}
# Load evals
var_cont <- readPNG(file.path(fp, "brt_var_cont.png"))

# Arrange grid
grid.arrange(rasterGrob(var_cont),
             ncol = 1)
```

**Figure 15.** Biomod BRT variable contributions on a monthly time scale.

#### RF variable contribution

```{r rf_biomod_var_cont, echo=FALSE}
# Load evals
var_cont <- readPNG(file.path(fp, "rf_var_cont.png"))

# Arrange grid
grid.arrange(rasterGrob(var_cont),
             ncol = 1)
```

**Figure 16.** Biomod RF variable contributions on a monthly time scale.

### Actual abundance vs. predicted probability of suitability

#### Biomod ensemble

For the biomod ensemble models, the logged actual abundance of ``r params$species`` was plotted against the predicted probability of suitability.  Above the threshold of ``r params$threshold``, the probability of suitability tends to increase.  Below the threshold, the probability of suitability tends to be zero.  


```{r ensemble_actual_vs_pred, echo=FALSE}

# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod_Climatologies", "Plots")
# Load filenames
files <- list.files(fp)

# Load region plots
biomod1 <- readPNG(file.path(fp, "biomod_abund_vs_pred_1.png"))
biomod2 <- readPNG(file.path(fp, "biomod_abund_vs_pred_2.png"))
biomod3 <- readPNG(file.path(fp, "biomod_abund_vs_pred_3.png"))
biomod4 <- readPNG(file.path(fp, "biomod_abund_vs_pred_4.png"))
biomod5 <- readPNG(file.path(fp, "biomod_abund_vs_pred_5.png"))
biomod6 <- readPNG(file.path(fp, "biomod_abund_vs_pred_6.png"))
biomod7 <- readPNG(file.path(fp, "biomod_abund_vs_pred_7.png"))
biomod8 <- readPNG(file.path(fp, "biomod_abund_vs_pred_8.png"))
biomod9 <- readPNG(file.path(fp, "biomod_abund_vs_pred_9.png"))
biomod10 <- readPNG(file.path(fp, "biomod_abund_vs_pred_10.png"))
biomod11 <- readPNG(file.path(fp, "biomod_abund_vs_pred_11.png"))
biomod12 <- readPNG(file.path(fp, "biomod_abund_vs_pred_12.png"))


# Arrange grid
grid.arrange(rasterGrob(biomod1), rasterGrob(biomod2), rasterGrob(biomod3),
             rasterGrob(biomod4), rasterGrob(biomod5), rasterGrob(biomod6),
             rasterGrob(biomod7), rasterGrob(biomod8), rasterGrob(biomod9),
             rasterGrob(biomod10), rasterGrob(biomod11), rasterGrob(biomod12),
             ncol=3)

```

**Figure 17.** Actual logged abundance versus predicted probability of suitability for ``r params$species`` for all 12 months.

### Ensemble region plots

```{r ensemble_region_plot, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_ensemble_abund_pred_monthly.png"))

if ("GBK_ensemble_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_ensemble_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_ensemble_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_ensemble_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

### GAM region plots

```{r biomod_gam_region_plot, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_gam_abund_pred_monthly.png"))

if ("GBK_gam_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_gam_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_gam_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_gam_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

### Ensemble region plots

```{r biomod_brt_region_plot, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_brt_abund_pred_monthly.png"))

if ("GBK_brt_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_brt_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_brt_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_brt_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

### Ensemble region plots

```{r biomod_rf_region_plot, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Biomod", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_rf_abund_pred_monthly.png"))

if ("GBK_rf_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_rf_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_rf_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_rf_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

## MGCV GAMs

### Climatology

The generalized additive models (GAMs) were run using the mgcv package and were used to model ``r params$species`` abundances.  

```{r display_GAM, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Plots")

# Load images
gam1 <- readPNG(file.path(fp, "gam_proj_1.png"))
gam2 <- readPNG(file.path(fp, "gam_proj_2.png"))
gam3 <- readPNG(file.path(fp, "gam_proj_3.png"))
gam4 <- readPNG(file.path(fp, "gam_proj_4.png"))
gam5 <- readPNG(file.path(fp, "gam_proj_5.png"))
gam6 <- readPNG(file.path(fp, "gam_proj_6.png"))
gam7 <- readPNG(file.path(fp, "gam_proj_7.png"))
gam8 <- readPNG(file.path(fp, "gam_proj_8.png"))
gam9 <- readPNG(file.path(fp, "gam_proj_9.png"))
gam10 <- readPNG(file.path(fp, "gam_proj_10.png"))
gam11 <- readPNG(file.path(fp, "gam_proj_11.png"))
gam12 <- readPNG(file.path(fp, "gam_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(gam1), rasterGrob(gam2), rasterGrob(gam3),
             rasterGrob(gam4), rasterGrob(gam5), rasterGrob(gam6),
             rasterGrob(gam7), rasterGrob(gam8), rasterGrob(gam9),
             rasterGrob(gam10), rasterGrob(gam11), rasterGrob(gam12),
             ncol=3)


```

**Figure 18.** Monthly climatological GAM projections.  The climatology was created by averaging together the projections from 2000 to 2017.

### Evaluations

```{r gam_evals, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Evals")

# Load evals
aic <- readPNG(file.path(fp, "gam_aic.png"))
rmse <- readPNG(file.path(fp, "gam_rmse.png"))
rsq <- readPNG(file.path(fp, "gam_rsq.png"))

# Arrange grid
grid.arrange(rasterGrob(aic), rasterGrob(rmse), rasterGrob(rsq),
             ncol=3)
```

**Figure 19.** Model evaluations on a monthly time scale using a.) AIC, b.) RMSE, and c.) R2.

### Regions

The study area was divided into three regions, the MAB, GBK, and GOM.  For each region, a climatological average (one point per month), monthly average time series, and annual average time series were computed and the actual versus predicted abundance values were plotted.

#### Climatological average

```{r gam_clim, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_gam_abund_pred_climatological.png"))

if ("GBK_gam_abund_pred_climatological.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_gam_abund_pred_climatological.png"))
}

gom <- readPNG(file.path(fp, "GOM_gam_abund_pred_climatological.png"))


# Arrange grid
if ("GBK_gam_abund_pred_climatological.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 20.** Climatological abundance values averaged over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

#### Monthly average time series

```{r gam_monthly, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "GAMs", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_gam_abund_pred_monthly.png"))

if ("GBK_gam_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_gam_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_gam_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_gam_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 21.** Abundance values averaged monthly over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

#### Annual average time series

```{r gam_annual, echo=FALSE}
# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_gam_abund_pred_annual.png"))

if ("GBK_gam_abund_pred_annual.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_gam_abund_pred_annual.png"))
}

gom <- readPNG(file.path(fp, "GOM_gam_abund_pred_annual.png"))


# Arrange grid
if ("GBK_gam_abund_pred_annual.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 22.** Abundance values averaged annually over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

### Actual vs predicted values

Actual abundance values were plotted against predicted abundance values.

```{r gam_actual_vs_pred, echo=FALSE}

# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Plots")

# Load region plots
gam1 <- readPNG(file.path(fp, "gam_abund_vs_pred_1.png"))
gam2 <- readPNG(file.path(fp, "gam_abund_vs_pred_2.png"))
gam3 <- readPNG(file.path(fp, "gam_abund_vs_pred_3.png"))
gam4 <- readPNG(file.path(fp, "gam_abund_vs_pred_4.png"))
gam5 <- readPNG(file.path(fp, "gam_abund_vs_pred_5.png"))
gam6 <- readPNG(file.path(fp, "gam_abund_vs_pred_6.png"))
gam7 <- readPNG(file.path(fp, "gam_abund_vs_pred_7.png"))
gam8 <- readPNG(file.path(fp, "gam_abund_vs_pred_8.png"))
gam9 <- readPNG(file.path(fp, "gam_abund_vs_pred_9.png"))
gam10 <- readPNG(file.path(fp, "gam_abund_vs_pred_10.png"))
gam11 <- readPNG(file.path(fp, "gam_abund_vs_pred_11.png"))
gam12 <- readPNG(file.path(fp, "gam_abund_vs_pred_12.png"))


# Arrange grid
grid.arrange(rasterGrob(gam1), rasterGrob(gam2), rasterGrob(gam3),
             rasterGrob(gam4), rasterGrob(gam5), rasterGrob(gam6),
             rasterGrob(gam7), rasterGrob(gam8), rasterGrob(gam9),
             rasterGrob(gam10), rasterGrob(gam11), rasterGrob(gam12),
             ncol=3)

```

**Figure 23.** Actual logged abundance versus logged predicted abundance for ``r params$species`` for all 12 months.

## GBM BRTs

### Climatology

The boosted regression tree (BRT) models were run using the gbm package and used to model ``r params$species`` abundances. 

```{r display_BRT, echo=FALSE}
# Load images
brt1 <- readPNG(file.path(fp, "brt_proj_1.png"))
brt2 <- readPNG(file.path(fp, "brt_proj_2.png"))
brt3 <- readPNG(file.path(fp, "brt_proj_3.png"))
brt4 <- readPNG(file.path(fp, "brt_proj_4.png"))
brt5 <- readPNG(file.path(fp, "brt_proj_5.png"))
brt6 <- readPNG(file.path(fp, "brt_proj_6.png"))
brt7 <- readPNG(file.path(fp, "brt_proj_7.png"))
brt8 <- readPNG(file.path(fp, "brt_proj_8.png"))
brt9 <- readPNG(file.path(fp, "brt_proj_9.png"))
brt10 <- readPNG(file.path(fp, "brt_proj_10.png"))
brt11 <- readPNG(file.path(fp, "brt_proj_11.png"))
brt12 <- readPNG(file.path(fp, "brt_proj_12.png"))

# Arrange grid
grid.arrange(rasterGrob(brt1), rasterGrob(brt2), rasterGrob(brt3),
             rasterGrob(brt4), rasterGrob(brt5), rasterGrob(brt6),
             rasterGrob(brt7), rasterGrob(brt8), rasterGrob(brt9),
             rasterGrob(brt10), rasterGrob(brt11), rasterGrob(brt12),
             ncol=3)
```

**Figure 24.** Monthly climatological BRT projections.  The climatology was created by averaging together the projections from 2000 to 2017.

### Evaluations

```{r brt_evals, echo=FALSE}
# Initialize file path
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Evals")

# Load evals
rmse <- readPNG(file.path(fp, "brt_rmse.png"))
rsq <- readPNG(file.path(fp, "brt_rsq.png"))

# Arrange grid
grid.arrange(rasterGrob(rmse), rasterGrob(rsq),
             ncol=2)
```

**Figure 25.** Model evaluations on a monthly time scale using a.) RMSE and b.) R2.

### Regions

The study area was divided into three regions, the MAB, GBK, and GOM.  For each region, a climatological average (one point per month), monthly average time series, and annual average time series were computed and the actual versus predicted abundance values were plotted.

#### Climatological average

```{r brt_clim, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_brt_abund_pred_climatological.png"))

if ("GBK_brt_abund_pred_climatological.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_brt_abund_pred_climatological.png"))
}

gom <- readPNG(file.path(fp, "GOM_brt_abund_pred_climatological.png"))


# Arrange grid
if ("GBK_brt_abund_pred_climatological.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 26.** Climatological abundance values averaged over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

#### Monthly average time series

```{r brt_monthly, echo=FALSE}
# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "BRTs", "Plots")

# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_brt_abund_pred_monthly.png"))

if ("GBK_brt_abund_pred_monthly.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_brt_abund_pred_monthly.png"))
}

gom <- readPNG(file.path(fp, "GOM_brt_abund_pred_monthly.png"))


# Arrange grid
if ("GBK_brt_abund_pred_monthly.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 27.** Abundance values averaged monthly over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

#### Annual average time series

```{r brt_annual, echo=FALSE}
# Load filenames
files <- list.files(fp)

# Load region plots
mab <- readPNG(file.path(fp, "MAB_brt_abund_pred_annual.png"))

if ("GBK_brt_abund_pred_annual.png" %in% files) {
  gbk <- readPNG(file.path(fp, "GBK_brt_abund_pred_annual.png"))
}

gom <- readPNG(file.path(fp, "GOM_brt_abund_pred_annual.png"))


# Arrange grid
if ("GBK_brt_abund_pred_annual.png" %in% files) {
  grid.arrange(rasterGrob(mab), rasterGrob(gbk), rasterGrob(gom),
              ncol=3)
} else {
  grid.arrange(rasterGrob(mab), rasterGrob(gom),
              ncol=2)
}

```

**Figure 28.** Abundance values averaged annually over three regions, a.) the Mid Atlantic Bight, b.) George's Bank, and c.) the Gulf of Maine.  The predicted values come from the projections.

### Actual vs. predicted values

Actual abundance values were plotted against predicted abundance values.


```{r brt_actual_vs_pred, echo=FALSE}

# Initialize filepath
fp <- file.path(params$fp_out, params$species, params$version, "Climatologies", "Plots")

# Load region plots
brt1 <- readPNG(file.path(fp, "brt_abund_vs_pred_1.png"))
brt2 <- readPNG(file.path(fp, "brt_abund_vs_pred_2.png"))
brt3 <- readPNG(file.path(fp, "brt_abund_vs_pred_3.png"))
brt4 <- readPNG(file.path(fp, "brt_abund_vs_pred_4.png"))
brt5 <- readPNG(file.path(fp, "brt_abund_vs_pred_5.png"))
brt6 <- readPNG(file.path(fp, "brt_abund_vs_pred_6.png"))
brt7 <- readPNG(file.path(fp, "brt_abund_vs_pred_7.png"))
brt8 <- readPNG(file.path(fp, "brt_abund_vs_pred_8.png"))
brt9 <- readPNG(file.path(fp, "brt_abund_vs_pred_9.png"))
brt10 <- readPNG(file.path(fp, "brt_abund_vs_pred_10.png"))
brt11 <- readPNG(file.path(fp, "brt_abund_vs_pred_11.png"))
brt12 <- readPNG(file.path(fp, "brt_abund_vs_pred_12.png"))


# Arrange grid
grid.arrange(rasterGrob(brt1), rasterGrob(brt2), rasterGrob(brt3),
             rasterGrob(brt4), rasterGrob(brt5), rasterGrob(brt6),
             rasterGrob(brt7), rasterGrob(brt8), rasterGrob(brt9),
             rasterGrob(brt10), rasterGrob(brt11), rasterGrob(brt12),
             ncol=3)

```

**Figure 29.** Actual logged abundance versus logged predicted abundance for ``r params$species`` for all 12 months.


